### 2018-04-23 QualifyingReport/QualifyingReport.tex
workplan updated

\documentclass{article}

\title{Qualifying Report}
\date{2018-04-22}
\author{Jomar Alcantara}

\begin{document}
	\maketitle
	\newpage
	\section{Overview of Problem}
	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment. As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language. I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.

	\section{Literature Review}

	\section{Work already carried out}

	\section{Work Plan for the future}

	\section{Proposed Timetable}
	\subsection{May 2019}

	\subsection{June 2019}

	\subsection{July 2019}

	\subsection{August 2019}

	\subsection{September 2019}

	\subsection{October 2019}

	\subsection{November 2019}

	\subsection{December 2019}

	\subsection{January 2020}

	\subsection{February 2020}

	\subsection{March 2020}

	\subsection{April 2020}

	\subsection{May 2020}

	\subsection{June 2020}

	\subsection{July 2020}

	\subsection{August 2020}

	\subsection{September 2020}

	\subsection{October 2020}

	\subsection{November 2020}

	\subsection{December 2020}

	\subsection{January 2021}

	\subsection{February 2021}

	\subsection{March 2021}

	\subsection{April 2021}
	\section{Provision Table of Contents}
\end{document}