### 2018-04-26 QualifyingReport/QualifyingReport.tex
continue to work on literature review and make amendments to other documents

\documentclass{article}

\title{Qualifying Report}
\date{2018-04-22}
\author{Jomar Alcantara}

\begin{document}
	\maketitle
	\newpage
	\section{Overview of Problem}
	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment. As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language. I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.

	\section{Literature Review}

	\section{Work already carried out}

	\subsection{April 2018}

	Thursday, 26th April - Workplan (First Draft Agreed)

	\section{Work Plan for the future}

	\section{Proposed Timetable}

	\subsection{May 2018}
	Wednesday, 9th May - Alexa Skills Training (All Day)
	\subsection{June 2018}
	Friday 15th June - First Draft of Literature Review to be handed in.
	\newline
	Monday 18th June - Monday 25th June - Annual Leave
	\subsection{July 2018}

	\subsection{August 2018}

	\subsection{September 2018}

	\subsection{October 2018}

	\subsection{November 2018}

	\subsection{December 2018}

	\subsection{January 2019}
	Friday 11th January - Rough Draft of Qualifying Due
	Friday 18th January - Complete first rough pass of capturing and processing speech.

	\subsection{February 2019}
	Friday, 1st February - Qualifying Report Due

	\subsection{March 2019}

	\subsection{April 2019}
	Monday, 1st April - Viva Due by this point.
	\subsection{May 2019}

	\subsection{June 2019}

	\subsection{July 2019}

	\subsection{August 2019}

	\subsection{September 2019}

	\subsection{October 2019}

	\subsection{November 2019}

	\subsection{December 2019}

	\subsection{January 2020}

	\subsection{February 2020}

	\subsection{March 2020}

	\subsection{April 2020}

	\subsection{May 2020}

	\subsection{June 2020}

	\subsection{July 2020}

	\subsection{August 2020}

	\subsection{September 2020}

	\subsection{October 2020}

	\subsection{November 2020}

	\subsection{December 2020}

	\subsection{January 2021}

	\subsection{February 2021}

	\subsection{March 2021}

	\subsection{April 2021}
	\section{Provision Table of Contents}
	\subsection{Chapter 1: Introduction}
	\subsection{Chapter 2: Literature Review}
	\subsection{Chapter 3: New theories, Description of experiments}
	\subsection{Chapter 4: Results}
	\subsection{Chapter 5: Discussion}
	\subsection{Chapter 6: Conclusion}
\end{document}