### 2018-04-26 Workplan/Workplan.tex

\documentclass{article}

\title{Workplan}
\date{2018-04-17}
\author{Jomar Alcantara}

\begin{document}
	\maketitle
	\section{Interest Statement}
	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment. As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language. I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.
	\section{Literature Review}
	\subsection{Dementia - Context}
	What is the Problem? 
	Why is this important?
	What is the potential impact?
	\subsection{How do Psychologists assess dementia}
	What tools do psychologists use to diagnose dementia? 
	Mini mental state examination (MMSE), 
	Addenbrooke's Cognitive Examination(ACE-III)
	Wechsler Adult Intelligence Scale(WAIS)
	The Free and Cued Selective Reminding Test (FCSRT)
	\subsection{How do Psychologists understand how language is affected by dementia}
	What language deficits are there and how are they characterised?
	What speech deficits are there and how are they characterised?
	\subsection{How is Natural Language Processing best positioned to help in this area?}
	\subsection{Are there any other technologies that can be used to help diagnosis?}

	\subsection{What Research has been done in the area so far?}
	\subsection{Conclusion}

	\section{Stages of research}
	\section{Stage 1 - Data capture}
	\subsection{Training needs analysis}
	\subsection{Psychology}
	Training in delivering the WAIS or psychological tools that might be needed 
	\subsection{Programming}
	\begin{itemize}
		\item Online courses in Python Programming
		\item Online courses in Machine Learning
	\end{itemize}

	\section{Data collection}
	\subsection{DementiaBank and AphasiaBank}
	\textbf{Holland Corpus}
	\newline 
	Consists of 2 participants a 62 year old man with mild AD and a 78 year old woman with more advanced AD.
	\newline
	\textbf{Kempler Corpus}
	\newline
	Consists of 6 participants with a range of 65 to 86 (mean, sd)
	\newline
	\textbf{Pitt Corpus}
	\newline
	Consists of 104 control participants, 208 with dementia and 85 with an unknown diagnosis.
	\newline
	\textbf{DePaul Corpus}
	\newline
	Consists of 1 participant, 66 year old female with PPA.
	\newline
	\textbf{Hopkins Corpus}
	\newline
	Consists of 36 participants, 23 males and 13 females

	\subsection {Three Authors, P.D. James, Iris Murdoch and Agatha Christie}	
	Data to be downloaded

	\subsection{Presidents Speeches}
	Data to be downloaded

	\subsection{Alexa Skills Kit and Amazon Web Services}
	\begin{itemize}
		\item \textbf{Training in Alexa Skills Kit - Booked}
		\item Training in Amazon Web Services
	\end{itemize}

	\section{Stage 2 - Data Processing, Feature Construction}
	\subsection{Natural Language Processing}
	Online courses in Natural Language Processing

	\begin{itemize}
		\item Jurafsky and Martin - Speech and Language Processing 
		\item Manning and Schutze - Statistical NLP 
		\item Manning and Raghavan - Intro to IR
	\end{itemize}

	\section{Stage 3 - Data Analysis and Machine Learning}
	\subsection{Application of Machine Learning Algorithms}

	\section{Proposed Timeline, Key Dates and Deliverables}
	\subsection{April 2018}
	Thursday, 26th April - Workplan (First Draft Agreed)
	\subsection{May 2018}
	Wednesday, 9th May - Alexa Skills Training (All Day)
	\subsection{June 2018}
	Friday 15th June - First Draft of Literature Review to be handed in.
	\newline
	Monday 18th June - Monday 25th June - Annual Leave
	\subsection{July 2018}

	\subsection{August 2018}
	Friday 10th August - First draft of experimental proposal to be submitted.

	\subsection{September 2018}
	Friday 21st September - Final draft of experimental proposal to be submitted. 
	Friday 21st September - NHS Ethics forms to be sorted  - Takes 60 days!
	\subsection{October 2018}

	\subsection{November 2018}

	\subsection{December 2018}

	\subsection{January 2019}
	Friday 11th January - Rough Draft of Qualifying Due
	Friday 18th January - Complete first rough pass of capturing and processing speech.

	\subsection{February 2019}
	Friday, 1st February - Qualifying Report Due

	\subsection{March 2018}

	\subsection{April 2018}
	Monday, 1st April - Viva Due by this point.
\end{document}