### 2018-07-09 ProjectPlan/ProjectPlan.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Class
\documentclass[a4paper]{article}
\title{Project Plan}
\date{2018-06-06}
\author{Jomar Alcantara}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin {document}
\section{Aims and Objectives}
\section{Tasks}
	\subsection{April 2018}
	Thursday, 26th April - Workplan (First Draft Agreed)
	\subsection{May 2018}
	Wednesday, 9th May - Alexa Skills Training (All Day)
	\subsection{June 2018}
	Friday 15th June - First Draft of Literature Review to be handed in.
	\newline
	Monday 18th June - Monday 25th June - Annual Leave
	\subsection{July 2018}
	Friday 13th July - Written Assignment (IS4001) Due
	Friday 13th July - Project Plan (IS4001) Due

	\subsection{August 2018}
	Friday 10th August - First draft of experimental proposal to be submitted.

	\subsection{September 2018}
	Friday 21st September - Final draft of experimental proposal to be submitted. 
	Friday 21st September - NHS Ethics forms to be sorted  - Takes 60 days!
	\subsection{October 2018}

	\subsection{November 2018}

	\subsection{December 2018}

	\subsection{January 2019}
	Friday 11th January - Rough Draft of Qualifying Due
	Friday 18th January - Complete first rough pass of capturing and processing speech.
	Thursday, 31st January - Qualifying Report Due

	\subsection{February 2019}

	\subsection{March 2019}

	\subsection{April 2019}
	Monday, 1st April - Viva Due by this point.
\section{Obstacles}
\end {document}