### 2018-07-13 ExperimentProposals/Experiments.tex
tidy up of folders and moved LR3K files to archive

\documentclass{article}

\title{Experimental Report}
\date{2018-04-22}
\author{Jomar Alcantara}

\begin{document}
	\maketitle

\bibliographystyle{unsrt}
\bibliography{ExperimentReport}

\end{document}