### 2018-07-19 ExperimentProposals/Experiments.tex
work done 19/07-18

\documentclass{article}

\title{Comparing the presidential news conferences of Ronald Reagan, George H.W. Bush and Donald Trump - A exploration of deep learning for the purposes of tracking language changes over time.\newline \\\large Experimental Proposal 1 }
\date{2018-07-19}
\author{Jomar Alcantara}

\begin{document}
\maketitle
\newpage
\tableofcontents
\newpage
\section{Abstract}
\section{Introduction}
\section{Methodology}
\section{Results}
\section{Discussion and Future Work}
\section{Conclusion}
\bibliographystyle{unsrt}
\bibliography{ExperimentReport}
\end{document}