### 2018-07-19 LiteratureReview/MeasuresOfLinguisticComplexity.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Class
\documentclass[a4paper]{article}
%% Packages
\input{settings/packages}
%% Page Settings
\input{settings/page}
%% Own Commands
\input{settings/macros}
\title{Measures of Linguistic Complexity: A literature Review}
\date{2018-07-13}
\author{Jomar Alcantara}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin {document}
\maketitle
\newpage
\tableofcontents
\newpage
\section{Introduction}
\section{Measures of Lexical Richness}
\subsection{Type token ratio(TTR)}
Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
\begin{equation} \label{x1}
    TTR = number of unique words / total number of words.
\end{equation}

\newline
\par
\subsection{Brunet's Index(W)}
\subsection{Honore's Scale}
\section{Conclusion}
\bibliographystyle{unsrt}
\bibliography{LiteratureReview}
\end {document}