### 2019-03-22 Archive/DhaneshaPersonSpec/PersonSpec.tex
reorganised files

\documentclass[11pt]{article}
\begin{document}
\textbf{MAIN DUTIES AND RESPONSIBILITIES}
\begin{enumerate}
	\item To support and contribute to the research activity of the project
	\item To collect data from project participants
	\item To recruit project participants from Join Dementia Research databases and other sources as guided and to identify needs and priorities
	\item To work independently according to a plan agreed with project staff
	\item To manage data for statistical analysis including data entry and data cleaning
	\item To ensure the secure and proper maintenance of relevant project files, research materials and data
	\item To assist in the preparation of interim and final reports for the project lead
	\item To be willing to travel as this project may involve working at other sites
	\item To assist the project team to ensure that all deadlines are met and that the study progresses according to the agreed timetable
	\item To seek and utilize supervision from the Project Lead as appropriate
	\item To carry out administrative duties required for co-ordination of the project
	\item To attend such meetings as necessary for satisfactory completion of the duties of the post
	\item Through liaison with clinical, administrative and managerial staff, to ensure that that ethical standards relating to confidentiality and the protection of information are adhered to
	\item Sensitively deal with emotive and distressing topics through project
	\item Sensitively and effectively work with patient groups to explore a range of emotive and highly contentious subjects
	\item Regularly undertake travel to a variety of sites using public transport
	\item Use Visual Display Unit equipment more or less continuously on almost all days
	\item To prioritise own workload, seeking guidance from and accepting the advice of the Project Lead where appropriate.  
	\item To be involved in the discussion of any queries associated with the project with the participants and professionals involved, in consultation with the project lead
	\item To conduct all other duties as agreed by the project lead 
\end{enumerate}

\end{document}