### 2019-04-04 Source/Systematic Review/Systematic Review - Protocol.tex
added Systematic Review

\documentclass[12pt]{article}
\begin{document}
% https://github.com/anadallora/SNAC/blob/master/protocol%20v4.pdf
\section{Introduction}
\section{Research Protocol}
\subsection{Research Question}
\subsection{Search Process}
\subsection{Primary Study Selection}
\subsection{Study Quality Assessment}
\subsection{Data Extraction}
\subsection{Data Synthesis and Reporting}
\end{document}