### 2019-08-22 Source/Thesis/Thesis.tex
updated Thesis

%%% PREAMBLE
\documentclass[10pt, letterpaper, twoside, openany]{thesis}

\usepackage{setspace}
\doublespacing

\usepackage[utf8]{inputenc}
\usepackage{fdsymbol} 
\usepackage{graphicx}
\usepackage{float}

\DeclareUnicodeCharacter{03B2}{\ensuremath{\beta}}

\title{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
\author{
        Jomar Alcantara \\
        Department of Computer Science \\
        School of Engineering and Applied Sciences \\
        Aston University\\
        Birmingham, United Kingdom
}
\date{\today}

\begin{document}

\maketitle
\newpage
% Abstract
\section*{Abstract}
My original contribution to knowledge is....
\newpage
% Acknowledgements
\section*{Acknowledgements}
Please note: An editor has not been used in the construction of this thesis.
\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables

\chapter{Introduction}
\include{Chapter1}

\chapter{Background and Related Work, Systematic Review of NLP and Machine Learning Research}
\include{Chapter2}

\chapter{Delphi Methodology and developing consensus on how best to collect language samples using technology}
\include{Chapter3}

\chapter{Development of a pipeline that processes language data accurately}
\include{Chapter4}

\chapter{Analysis of DementiaBank dataset}
\include{Chapter5}

\chapter{Pilot study of the methodology developed}
\include{Chapter6}

\chapter{General Discussion, Conclusions and Future Work}
\include{Chapter7}

\bibliographystyle{unsrt}
\bibliography{main.bib}

\end{document}